### 2015-12-08 abstract.tex
Added one sentence to the abstract.
Fix abstract
Added sentence explaining lower bounds for mirror decsent.
Reworked introduction

We design algorithms for online linear optimization that have optimal regret
and at the same time do not need to know any upper or lower bounds on the norm
of the loss vectors.  We achieve adaptiveness to norms of loss vectors by scale
invariance, i.e., our algorithms make exactly the same decisions if the
sequence of loss vectors is multiplied by any positive constant.  Our
algorithms work for any decision set, bounded or unbounded.  For unbounded
decisions sets, these are the first truly adaptive algorithms for online linear
optimization.

>>>>

We design algorithms for online linear optimization that have optimal regret
and at the same time do not need to know any upper or lower bounds on the norm
of the loss vectors.  We achieve adaptiveness to norms of loss vectors by scale
invariance, i.e., our algorithms make exactly the same decisions if the
sequence of loss vectors is multiplied by any positive constant.  Our
algorithms work for any decision set, bounded or unbounded.  For unbounded
decisions sets, these are the first adaptive algorithms for online linear
optimization.

We also study a popular scale-free variant of online mirror descent algorithm,
and we show that in two natural settings it has linear or worse regret.

### 2015-12-08 introduction.tex
Reworked introduction

The contribution of this paper is twofold. First, in
Section~\ref{section:ada-ftrl} we show that the analysis and design of
\textsc{AdaHedge} can be generalized to the OLO scenario and to any strongly
convex regularizer, in an algorithm we call \textsc{AdaFTRL}, providing a new
and rather interesting way to adapt the learning rates to have scale-free
algorithms.  Second, in Section~\ref{section:solo-ftrl} we propose a new and
simple algorithm, \textsc{SOLO FTRL}, that is scale-free and is the
\textbf{first} scale-free online algorithm for unbounded sets with a
non-vacuous regret bound.  Both algorithms are instances of \textsc{Follow The
Regularized Leader} (\textsc{FTRL}) with an adaptive learning rate.  Moreover,
our algorithms show that scale-free algorithms can be obtained in a ``native''
and simple way, i.e.  without using ``doubling tricks'' that attempt to fix
poorly designed algorithms rather than directly solving the problem.

For both algorithms, we prove that for bounded decision sets the regret after
$T$ rounds is at most $O(\sqrt{\sum_{t=1}^T\|\ell_t\|_*^2})$.  We show that the
$\sqrt{\sum_{t=1}^T \|\ell_t\|_*^2}$ term is necessary by proving a $\Omega (D
\sqrt{\sum_{t=1}^T\|\ell_t\|_*^2} )$ lower bound on the regret of any algorithm
for OLO for any decision set with diameter $D$ with respect to the primal norm
$\|\cdot\|$. For the \textsc{SOLO FTRL} algorithm, we prove an
$O(\max_{t=1,2,\dots,T} \|\ell_t\|_* \sqrt{T})$ regret bound for any unbounded
decision set.

Our algorithms are also \textbf{any-time}, i.e., do not need to know the
number of rounds in advance and our regret bounds hold for all time steps
simultaneously.

>>>>

We design and analyze three scale-free algorithms. First two algorithms, which
we call \textsc{AdaFTRL} and \textsc{SOLO FTRL}, are instances of
\textsc{Follow The Regularized Leader} (\textsc{FTRL}) with adaptive learning
rate.  \textsc{AdaFTRL} is a generalization of
\textsc{AdaHedge}~\cite{de-Rooij-van-Erven-Grunwald-Koolen-2014} to arbitrary
strongly convex regularizers.  \textsc{SOLO FTRL} can be viewed as the
``correct'' scale-free version of \textsc{AdaGrad
FTRL}~\cite{Duchi-Hazan-Singer-2011} generalized to arbitrary strongly convex
regularizers.  The third algorithm, which we call \textsc{Scale-Free Mirror
Descent}, is a variant of \textsc{Mirror Descent}. It is a generalization of
\textsc{AdaGrad MD}~\cite{Duchi-Hazan-Singer-2011} to arbitrary strongly convex
regularizers, similar to \textsc{Optimistic MD}~\cite{Rakhlin-Sridharan-2013}.
The three algorithms are presented and analyzed in
Sections~\ref{section:ada-ftrl}, \ref{section:solo-ftrl} and
\ref{section:mirror-descent} respectively.

For all three algorithms, we prove that for bounded decision sets the regret
after $T$ rounds is at most $O \left(\sqrt{\sum_{t=1}^T\|\ell_t\|_*^2} \right)$
where the constant hidden in $O(\cdot)$ depends on the decision set and the
regularizer.  In Section~\ref{section:lower-bound}, we show that the
$\sqrt{\sum_{t=1}^T \|\ell_t\|_*^2}$ term is necessary by proving $\Omega
\left(D \sqrt{\sum_{t=1}^T\|\ell_t\|_*^2} \right)$ lower bound on the regret of
any algorithm for OLO for any decision set with diameter $D$ with respect to
the primal norm $\|\cdot\|$.

For the \textsc{SOLO FTRL} algorithm, we prove an $O \left(\max_{t=1,2,\dots,T}
\|\ell_t\|_* \sqrt{T} \right)$ regret bound for any unbounded decision set. The
constant hidden in $O(\cdot)$ depends not only on the decision set and the
regularizer, but also on the competitor $u \in K$.~\footnote{An upper bound
that does not depend on $u$ is impossible. It would contradict the $\Omega
\left( D \sqrt{\sum_{t=1}^T\|\ell_t\|_*^2} \right)$ lower bound.} This is the
\textbf{first adaptive algorithm for unbounded decision sets} with a
non-trivial upper bound.

All three algorithms are \textbf{any-time}, i.e., they do not need to know the
number of rounds in advance and the regret bounds hold for all time steps
simultaneously.

Our proofs of the regret upper bounds rely on homogeneous inequalities
(Lemmas~\ref{lemma:recurrence-solution},
\ref{lemma:sum-of-square-roots-inverses}, \ref{lemma:useful}).  Two of the
inequalities are new. Previously, many proofs of regret bounds relied on ad-hoc
non-homogeneous inequalities, or the existing algorithms used quantities (e.g.
$\sum_{t=1}^T \|\ell_t\|_*^2$) that are not available until all loss vectors
are processed.

Finally, in Section~\ref{subsection:mirror-descent-lower-bound}, we show
negative results for existing popular variants of \textsc{Mirror Descent}. We
show that variants of \textsc{Mirror Descent} have $\Omega(T)$ regret if the
Bregman divergence associated with the regularizer is unbounded on the decision
set.  The setting includes the standard gradient descent with step size
$1/\sqrt{t}$, \textsc{AdaGrad MD} and \textsc{Scale-Free Mirror Descent} on a
unbounded decision set. This result suggests that \textsc{Follow The
Regularized Leader} is superior to \textsc{Mirror Descent}.
